---
title: "Analyze and improve your data analysis workflow"
description: |
  A concise guide to R package drake. 
categories: 
  - data science
  - R
author:
  - name: Miha Gazvoda
    url: https://mihagazvoda.com
date: 04-21-2020
preview: https://camo.githubusercontent.com/44749362ca36c9e3295f2bcf18975d811564c121/68747470733a2f2f646f63732e726f70656e7363692e6f72672f6472616b652f7265666572656e63652f666967757265732f6c6f676f2e737667
output:
  distill::distill_article:
    self_contained: false
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

[`drake`](https://github.com/ropensci/drake) is an R package by [Will Landau](https://wlandau.github.io/) that analyzes your workflow. It

* enables to skip steps of the analysis with up-to-date results;  
* provides evidence that results match the underlying code and data;
* encourages good programming practices by modularizing your code into functions^[[More about functions in R.](https://r4ds.had.co.nz/functions.html)];
* interactively visualizes network representation of your workflow. 

<!-- ![](https://camo.githubusercontent.com/04d35d8cc5a548445bfc33840a88289fdc340ca3/68747470733a2f2f646f63732e726f70656e7363692e6f72672f6472616b652f7265666572656e63652f666967757265732f696e666f677261706869632e737667) -->

<!-- TODO include image that: X 1. Launch the code. 2. Wait while it runs. 3. Discover an issue. 4. Restart from scratch. -->

# Setup
Install `drake`. You can also load an example written by [Kirill Müller](https://github.com/krlmlr). It will appear in a new `main` folder. I will use it as a showcase for some file examples.  

```{r, eval=F, echo=T}
# Install and load drake
install.packages("drake")

# Get an example in a new `main` folder
drake::drake_example("main")

# You can use drake::examples() to see all examples
```

## Project structure
It's suggested that you start your project using this structure^[It's almost the same as in the example.]: 

```markdown
make.R
R/
├── packages.R
├── functions.R
└── plan.R
data/
```
You can also use `dflow::use_dflow()` to create almost similar structure. 

### Make
`make.R` is a master script that 

* loads your packages and functions^[If there are many functions, split them up into multiple files.];
* creates a `drake` plan;
* calls `make()`.

```{r, eval=F, echo=T}
# make.R
source("R/packages.R")  # loads packages
source("R/functions.R") # loads user-defined functions
source("R/plan.R")      # creates drake plan

make(plan)              # defined in R/plan.R
```

### Plan

`drake` plan is the high-level catalog of data analysis steps (such as data cleaning, model fitting, visualization, and reporting) in a workflow.  
Plan is presented as a data frame with columns named `target` and `command`.

```{r, echo = TRUE}
# plan.R
# The workflow `plan` data frame outlines what you are going to do.
plan <- drake::drake_plan(
  # target, command
  raw_data = readxl::read_excel(file_in("raw_data.xlsx")),
  data = raw_data %>%
    mutate(Species = forcats::fct_inorder(Species)),
  hist = create_plot(data),
  fit = lm(Sepal.Width ~ Petal.Width + Species, data),
  report = rmarkdown::render(
    knitr_in("report.Rmd"),
    output_file = file_out("report.html"),
    quiet = TRUE
  )
)
```

Drake plan is presented as a data frame with columns named `target` and `command`. Each row represents a step in the workflow. Each command is a concise expression that makes use of our functions, and each target is the return value of the command. (The target column has the names of the targets, not the values.)

<details>
  <summary><b>See `plan` object.</b></summary>
```{r}
knitr::kable(plan)
```
</details>

<details>
  <summary><b>See dependency graph.</b></summary>
```{r}
drake::vis_drake_graph(plan)
```
</details>

<!-- TODO also include dependency plot -->

#### Choose good targets
As Will Landau proposed, a good target is

* long enough to eat up a decent chunk of runtime;
* small enough that make() frequently skips it;
* meaningful to your project;
* a well-behaved R object compatible with `saveRDS()`.


# Workflow
Even if you use `drake`, it makes sense to develop interactively. With `r_make("make.R")`^[If name make file `_drake.R` instead, then you are able to call `r_make()` without any argument.] you build your project. With `loadd` or `readd` you can return targets to your session and interactively use them to develop things further. 
<!-- TODO output? -->

## Basic commands

Here are the most useful commands.

```{r}
library(dplyr)
tibble::tibble(
  `function` = c("r_make()", "vis_drake_graph()", "code_to_function()", "loadd()", "readd()"),
  description = c("Build your project.", "Show an interactive visual network representation of your workflow.", "Create drake_plan()-ready functions from scripts.", "Loads built target(s) into your R session.", "Read and return a built target.")
  ) %>% 
knitr::kable()
```

You can find more functions in `drake` [README](https://github.com/ropensci/drake#function-reference).

<center>$\bullet\bullet\bullet$</center>

For further reading I suggest you [The drake R package User Manual](https://books.ropensci.org/drake/). The book also served as a resource for this post. 